Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-08-08 UMich/chapters/ch04-NetDelibExp.tex
More ch04 revisions.

Deliberation is fundamentally a social influence process in which
individuals exchange statements with their peers in an effort to influence their preferences.
We will thus benefit from a formalism to quantify and analzye social influence.
If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
Namely, we will construct measures of social influence based on the assumptions that:
\begin{itemize}
    \item The magnitude of an individual's change in preference increases with their susceptibility to social influence;
    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
\end{itemize}
The distance between two individuals' preferences is symmetric between the two individuals,
so the latter of the above conditions does not speak to who influences whom.
The converse of the former condition however suggests that the greater the change in an individual's preference,
the more they have been influenced by other.
Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
On the other hand, if the distance increases, we refer to this as negative social influence, suggesting one individual has influenced another to adopt an opposing preference.
We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.

@@

Our results show that in our experiment, the random-pod group 
experienced a small but robust increase in agreement,
wile the control group experienced a sharp increase followed by an even sharper decrease.
These results support the hypothesis H1 that network deliberation contributes to increased agreement relative to conventional single-group deliberation.
However, this effect is due more to a decrease in agreement
within the control group than an increase in the random-pod group.
After an initial increase in agreement, both groups showed an unexpected ``rebound'' effect.
In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.
In other words, participants altered their preferences to more strongly oppose those they interacted with.

@@

We hypothesized (H1) that network deliberation tends to reach a higher level of agreement than conventional single-group deliberation.
The observed agreement calculated from polls throughout the experiment experiment supports this hypothesis (Section \ref{sec:res-agreement}).
However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
leading us to expect a large increase in agreement in the random-pod group.
Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.
So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.

@@

>>>>

Deliberation is fundamentally a social influence process in which
individuals exchange statements with their peers in an effort to influence their preferences.
We will thus benefit from a formalism to quantify and analzye social influence.
If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
Namely, we will construct measures of social influence based on the assumptions that:
\begin{itemize}
    \item The magnitude of an individual's change in preference increases with their susceptibility to social influence;
    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
\end{itemize}
The distance between two individuals' preferences is symmetric between the two individuals,
so the latter of the above conditions does not speak to who influences whom.
The converse of the former condition however suggests that the greater the change in an individual's preference,
the more they have been influenced by other.
Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
On the other hand, if the distance increases, we refer to this as negative social influence, suggesting one individual has influenced another to adopt an opposing preference.
Negative social influence has the potential to produce polarization when individuals with similar views
remain similar to each other, but adjust their views to more strongly oppose the rest of the group.
We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.

@@

Our results show that in our experiment, the random-pod group 
experienced a small but robust increase in agreement,
wile the control group experienced a sharp increase followed by an even sharper decrease.
These results support the hypothesis H1 that network deliberation contributes to increased agreement relative to conventional single-group deliberation.
However, this effect is due more to a decrease in agreement
within the control group than an increase in the random-pod group.
After an initial increase in agreement, both groups showed an unexpected ``rebound'' effect.
In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.
In other words, participants altered their preferences to more strongly oppose those they interacted with.
Negative social influence is of interest as a potential driver of polarization.

@@

We hypothesized (H1) that network deliberation tends to reach a higher level of agreement than conventional single-group deliberation.
The observed agreement calculated from polls throughout the experiment experiment supports this hypothesis (Section \ref{sec:res-agreement}).
However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
leading us to expect a large increase in agreement in the random-pod group.
Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.
So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.
The participants began with a relatively high level of agreement and disagreements were minor, so a more controversial topic may be necessary to identify potential conflict-resolution mechanisms.

@@

While our experiment included both a control and experimental condition, having only one of each is a potential limitation.
To veryify the robustness of our results, future work incorporating multiple control and treatment groups is needed.
